---
title: "Reproducable Research Week2 Assignment"
author: "Veera"
date: "August 16, 2018"
output:
  html_document: default
  pdf_document: default
---

# Reproducable Research Week1 Assignment 
***
##Synospsis
***
 Based on the data from a personal activity monitoring device, adressing the below four questions
 
1. What is mean total number of steps taken per day?
  + Calculate the total number of steps taken per day
  + If you do not understand the difference between a histogram and a barplot, research the difference    between them. Make a histogram of the total number of steps taken each day
  + Calculate and report the mean and median of the total number of steps taken per day
2. What is the average daily activity pattern?
  + Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
  + Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
  
Download data from [here](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) and save data into one folder and set that  same folder as working directory.

# Load the data and see the Summary
***
```{r}
setwd(choose.dir())
Activity_data <- read.csv("activity.csv")
summary(Activity_data)
```

# Data Cleanup 
***
```{r}
Activity_data_set<-na.omit(Activity_data)
Activity_data_set$date<-as.Date(as.character(Activity_data_set$date))
```

# 1.What is mean total number of steps taken per day?
***
## 1.1 Calculate the total number of steps taken per day
```{r}
library(dplyr)
day_Activity_data<-Activity_data_set %>%
    group_by(date)  %>%
    summarise(total_steps=sum(steps,na.rm=TRUE))

```
First 10 Records are 
```{r}
head(day_Activity_data,10)
```
## 1.2 Histogram of the total number of steps taken each day
```{r}
library(plotly)
p <- plot_ly(x = day_Activity_data$total_steps , type = "histogram")
p
```

## 1.3 Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean_steps<-paste('Mean of total steps per day is mean_steps is ',mean(day_Activity_data$total_steps))
mean_steps
median_steps<-paste('Median of total steps per day is mean_steps is  ',median(day_Activity_data$total_steps))
median_steps
```
#2. What is the average daily activity pattern?
***
##2.1 Time Series Plot
```{r}

interval_average <- Activity_data %>% group_by(interval) %>% summarise(average = mean(steps,na.rm = TRUE))

p <- plot_ly(x = ~interval_average$interval, y = ~interval_average$average, mode = 'lines', text ="Average Daily Activity Pattern" )
p

```

##2.2 Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
maxIntervalsteps <- Activity_data[which.max(Activity_data$steps),]
maxIntervalsteps
```


#Imputing Missing Values 
```{r}
total_missed_data<-sum(rowSums(is.na(Activity_data)))
paste(total_missed_data,"records data is null")
```

##1.Strategy to fill missing vaues with mean
```{r}
mean_steps<-mean(Activity_data$steps,na.rm = TRUE)

```
##2.Strategy to fill missing vaues is Mean
```{r}

Activity_data[is.na(Activity_data$steps),"steps"]<-mean_steps
```
##3.Histogram After Impute
```{r}
library(plotly)
day_Activity_data_after_impute<-Activity_data %>%
    group_by(date)  %>%
    summarise(total_steps=sum(steps,na.rm=TRUE))
    
p <- plot_ly(x = day_Activity_data_after_impute$total_steps , type = "histogram")
p
```

#4.Are there differences in activity patterns between weekdays and weekends?
```{r}

Activity_data2<-Activity_data
Activity_data2$day<-weekdays(as.Date(Activity_data2$date))
Activity_data2$weektype <-"WD"
Activity_data2[Activity_data2$day %in% c('Sunday','Saturday'),"weektype" ]<- "WE"
interval_average2 <- Activity_data2 %>% group_by(interval,weektype) %>% summarise(average = mean(steps,na.rm = TRUE))
week_day_count<-interval_average2[interval_average2$weektype=="WD",]
week_end_count<-interval_average2[interval_average2$weektype=="WE",]
p <- plot_ly(week_day_count , x = ~interval, y = ~average, name = 'trace 0',  mode = 'scatter',text="Week Day Activity Report", line = list(color = 'red' ) )
q <- plot_ly(week_end_count , x = ~interval, y = ~average, name = 'trace 0',  mode = 'scatter' ,text="Week End Activity Report",line = list(color = 'orange' )) 
par(mfrow=c(1,2)) 
p
q



```

